Testable normative theories

### 2020-04-30 simple.tex
New conditions that hopefully wake sense

A normative theory is a tuple $(\allprops, ¬, \allargs, I, (\ileadstost[i])_{i \in I}, q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats)$,
with $\gargs \subseteq \allargs$, ${\ileadstost[i]} \subseteq (\allargs × \allargs) × \Phi$, ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
where $\Phi = \allprops × \set{\text{possible}, \text{sure}}$.

@@

As usual, $(\ar', \ar) \nileadstost \phi$ means $¬[(\ar', \ar) \ileadstost \phi]$.

Abusing notation, given $\ar \in \allargs$ and $\phi \in \Phi$, I write $\ar \ileadstost \phi$ to mean $[\forall \ar' \in \allargs: (\ar', \ar) \ileadstost \phi]$.

The set $\iphi \subseteq \Phi$ is the set of interest. 
When $(\prop, \text{possible}) \in \iphi$, $\prop$ is considered as a possibility in the deliberated judgment of $i$ (it is not rejected), and when $(\prop, \text{sure}) \in \iPhi$, $\prop$ is considered as a certainty in the deliberated judgment of $i$. TODO we should have $(\prop, \text{sure}) \in \iPhi ⇒ (¬\prop, \text{possible}) \notin \iPhi$.
Part of the content of $\Phi$ is known as a result of a correct normative and empirical analysis.

@@

Given $\phi \in {\gleadsto}(\allargs)$, an argument $\ar \in \allargs$ is \emph{in favor of $\phi$}, $\ar \in \gargs[\phi]$, iff $\ar \in [\gbeatsinv]^{2k}({\gleadstoinv}(\phi))$ for some $k \in \N$, where $R^\alpha$ denotes $\alpha$ times the composition of the relation $R$ with itself. (Throughout the article, $\N$ includes zero.) Define $\gargs = \cup_{\phi \in {\gleadstoinv}(\allargs)} \gargs[\phi]$.

Define ${\ibeatse} \subseteq \allargs × \gargs$ such that $\ar' \ibeatse \ar$ iff $\exists \phi \suchthat \ar \in \gargs[\phi] \text{ and } (\ar', \ar) \ileadstoe ¬\phi$. 
Given $\phi \in {\gleadsto}(\allargs)$, define ${\ibeatse[!\phi]} \subseteq \allargs × \gargs[\phi]$ such that $\ar' \ibeatse[!\phi] \ar$ iff $(\ar', \ar) \ileadstoe ¬\phi$. Define ${\ibeatse[\phi]} \subseteq \gargs[\phi] × \allargs$ such that $\ar' \ibeatse[\phi] \ar$ iff $(\ar', \ar) \ileadstoe \phi$. 
Define ${\ibeatsst} \subseteq \gargs × \allargs$ such that $\ar' \ibeatsst \ar$ iff $\forall \phi \suchthat \ar' \in \gargs[\phi]: (\ar', \ar) \ileadstost \phi$.
\footnote{When both $\ar_1, \ar_2 \in \gargs$, the relationship is as follows. $\ar_2 \ibeatse \ar_1 ⇒ \ar_2 \nibeatsst \ar_1$, the converse failing only if $\ar_2$ has some claims that $\ar_1$ does not have (thus $\exists \phi \suchthat \ar_2 \in \gargs[\phi] \land \ar_1 \notin \gargs[\phi]$): $\ar_2 \nibeatsst \ar_1$ exactly when $\ar_2$ fails to argue decisively for some of its claims, and $\ar_2 \nibeatse \ar_1$ exactly when $\ar_1$ argues decisively for all of its claims.}

A normative theory starts with an axiom stating its normative adequacy.
\begin{axiom}[Normative adequacy of $\allargs$ and $\ileadstost$]
	A general normative theory is \emph{normatively adequate} iff $\forall \phi \in \Phi$: 
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ 
		\begin{cases}
			\phi \in \iphi,\\
			\lnot\phi \notin \iphi.
		\end{cases}
	\end{equation}
\end{axiom}

@@

\begin{definition}[Validity]
	A theory is \emph{valid} iff ${\gleadsto}(\allargs) \subseteq \iphi$.
\end{definition}

The following condition mandates that if the arguments used by the theory are not decisive, it knows a reason.
\begin{definition}[Answerability]
	$\forall \phi \in \Phi, \ar_1 \in \allargs, \ar_2 \in \gargs[\phi] \cap {\gbeatsinv}(\ar_1): (\ar_2, \ar_1) \ileadstoe ¬\phi ⇒ [\exists \ar_3 \in {\gbeatsinv}(\ar_2) \suchthat (\ar_3, \ar_2) \ileadstoe \phi]$.
\end{definition}

\begin{definition}[Bounded width]
	$\forall \ar \in \gargs: \card{\ibeatseinv(\ar)} ≤ b_w$.
\end{definition}

\begin{definition}[Bounded length]
	$\forall n \in \N$, if $n > b_l$, there is no sequence of arguments $(\ar_k)_{0 ≤ k < n}$ such that $[\exists \phi \suchthat \forall (0 ≤ k < 2n): \ar_{2k} \in \gargs[\phi]]$ and $(\ar_{k + 1}, \ar_k) \ileadstoe \phi$ and $(\ar_{k + 1}, \ar_k) \ileadstoe ¬\phi$. \commentOC{This requires that $\ar_4$ have no attacker, not only in the sense that $\ar_4$ argues successfully against $\ar_3$, but in the (too strong) sense that $\ar_4$ argues successfully for $\phi$. It makes $\ar_4$ decisive, thus, requires that $\ar_4$ also argues against $\ar_1$, $\ar'_1, …$, and it makes the other arguments non useful. (Though may be useful for other dms, as we do not know in advance which one will be decisive.) I suppose I could simplify the presentation by making $\gargs[2]$ contain every good arguments and using only this one.}
\end{definition}

\begin{definition}[Reinstatement]
	A theory satisfies \emph{reinstatement} iff $\forall \ar_1, \ar_3 \in \gargs[\phi], \ar_2 \in \allargs: [\ar_1 \ileadstoe \phi, (\ar_2, \ar_1) \ileadstoe ¬\phi, (\ar_3, \ar_2) \ileadstoe \phi] ⇒ (\ar_1 + \ar_3, \ar_2) \ileadstoe \phi \land [\forall \ar_4 \ibeatse \ar_1 + \ar_3: \ar_4 \ibeatse \ar_1 \lor \ar_4 \ibeatse \ar_3]$.
\end{definition}

\begin{definition}[Operational validity]
	A theory is \emph{operationally valid} iff $\forall \phi \in {\gleadsto}(\allargs), \ar_0 \in \gargs[\phi], \ar_1 \in \allargs: (\ar_1, \ar_0) \ileadstoe ¬\phi ⇒ [\ar_1 \gbeats \ar_0] \land [\exists \ar_2 \in {\gbeatsinv}(\ar_1) \suchthat (\ar_2, \ar_1) \ileadstoe \phi]$.
\end{definition}

\begin{theorem}
	If a theory is answerable, satisfies reinstatement, has bounded width and length, and is operationally valid, it is valid. \commentOC{We don’t seem to need most of these conditions!}
\end{theorem}
\begin{proof}
	Given $\phi \in {\gleadsto}(\allargs), \args_2, \args_0 \subseteq \gargs$, define $\args_2$ as \emph{$\phi$-defending} $\args_0$ iff $\forall \ar_1 \in \allargs: [\exists \ar_0 \in \args_0 \suchthat (\ar_1, \ar_0) \ileadstoe ¬\phi] ⇒ [\exists \ar_2 \in \args_2 \suchthat (\ar_2, \ar_1) \ileadstoe \phi]$. Define $\args_2$ as \emph{decisively $\phi$-defending} $\args_0$ iff $\forall \ar_1 \in \allargs: [\exists \ar_0 \in \args_0 \suchthat (\ar_1, \ar_0) \ileadstoe ¬\phi] ⇒ [\exists \ar_2 \in \args_2 \suchthat (\ar_2, \ar_1) \ileadstost \phi]$.

>>>>

\subsection{Definition}
A normative theory is a tuple $(\allprops, ¬, \allargs, I, (\ileadstost[i])_{i \in I}, q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\ileadstost[i]} \subseteq (\allargs × \allargs) × \Phi$, ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
where $\Phi = \allprops × \set{\text{possible}, \text{sure}}$, and $+$ a commutative associative internal binary operation on $\allargs$. The notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument.

@@

As usual with binary relations, $(\ar', \ar) \nileadstost \phi$ means $¬[(\ar', \ar) \ileadstost \phi]$.
Abusing notation, given $\ar \in \allargs$ and $\phi \in \Phi$, I write $\ar \ileadstost \phi$ to mean $[\forall \ar' \in \allargs: (\ar', \ar) \ileadstost \phi]$.

The set $\iphi \subseteq \Phi$ is the set of interest. 
When $(\prop, \text{possible}) \in \iphi$, $\prop$ is considered as a possibility in the deliberated judgment of $i$ (it is not rejected), and when $(\prop, \text{sure}) \in \iPhi$, $\prop$ is considered as a certainty in the deliberated judgment of $i$. 
Part of the content of $\Phi$ is known as a result of a correct normative and empirical analysis.

Given $\ar_0 \in \gargs$, let $c(\ar_0) = \bigcup \set{\args \subseteq \gargs \suchthat \ar_0 = \sum \args} \subseteq \allargs$ denote the \emph{components} of $\ar_0$.

Given $\phi \in {\gleadsto}(\allargs)$, an argument $\ar \in \allargs$ is \emph{in favor of $\phi$}, $\ar \in \gargs[\phi, 0]$, iff $\ar \in [\gbeatsinv]^{2k}({\gleadstoinv}(\phi))$ for some $k \in \N$, where $R^\alpha$ denotes $\alpha$ times the composition of the relation $R$ with itself. (Throughout the article, $\N$ includes zero.) 
Define $\gargs[\phi] = \set{\ar \suchthat c(\ar) \subseteq \gargs[\phi, 0]}$.
Define $\gargs = \cup_{\phi \in {\gleadstoinv}(\allargs)} \gargs[\phi]$.

Given $\ar_2 \in \gargs, \ar_1 \in \allargs$, $\ar_2 \mbeats \ar_1$ has the semantic that $\gamma$ claims that $\ar_2$ trumps $\ar_1$ if necessary.
Given $\ar_1 \in \allargs, \ar_0 \in \gargs$, $\ar_1 \mbeats \ar_0$ has the semantic that $\gamma$ claims that it knows how to resist an attack from $\ar_1$ coming in response to using $\ar_0$, if it does come.

@@

Given $\phi \in {\gleadsto}(\allargs)$, define ${\ibeatse[!\phi]} \subseteq \allargs × \gargs[\phi]$ such that $\ar' \ibeatse[!\phi] \ar$ iff $(\ar', \ar) \ileadstoe ¬\phi$. Define ${\ibeatse[\phi]} \subseteq \gargs[\phi] × \allargs$ such that $\ar' \ibeatse[\phi] \ar$ iff $(\ar', \ar) \ileadstoe \phi$.

\subsection{Classify propositions}
A normative theory starts with an axiom stating its normative adequacy.
\begin{axiom}[Normative adequacy of $\allargs$ and $\ileadstost$]
	A general normative theory is \emph{normatively adequate} iff $\forall \phi \in \Phi$: 
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ 
		\begin{cases}
			\phi \in \iphi,\\
			\lnot\phi \notin \iphi.
		\end{cases}
	\end{equation}
\end{axiom}

@@

\subsection{Validity and conditions}
\begin{definition}[Validity]
	A theory is \emph{valid} iff ${\gleadsto}(\allargs) \subseteq \iphi$.
\end{definition}

$\ar_0 + \ar_2$ anticipates attacks from those attacking either, minus those attacked by any. (Actually any attacked may still attack the addition, but that’s because of some other not-yet-attacked, anticipated, attack.)

Define $\mbeats[g] \subseteq \gargs × \allargs$ as $\ar_2 \mbeats[g] \ar_1$ iff $\exists \ar \in c(\ar_2) \suchthat \ar \mbeats \ar_1$.
Define $\mbeats[a] \subseteq \allargs × \gargs$ as $\ar_1 \mbeats[a] \ar_0$ iff $\exists \ar \in c(\ar_0) \suchthat \ar_1 \mbeats \ar$ and $\nexists \ar \in c(\ar_0) \suchthat \ar \mbeats \ar_1]$. Define $\mbeats[*] \subseteq \allargs × \allargs$ as $\gbeatsstar = {\mbeats[g]} \cup {\mbeats[a]}$.

The following condition mandates that if the arguments used by the theory are not decisive, it knows a reason.
\begin{definition}[Answerability]
	$\forall \phi \in \Phi, \ar_1 \in \allargs, \ar_2 \in \gargs[\phi] \cap {\gbeatsstarinv}(\ar_1): (\ar_2, \ar_1) \ileadstoe ¬\phi ⇒ [\exists \ar_3 \in {\gbeatsstarinv}(\ar_2) \suchthat (\ar_3, \ar_2) \ileadstoe \phi]$.
\end{definition}

\begin{definition}[Bounded length]
	$\gargs$ must be finite.
\end{definition}

\begin{definition}[Operational validity]
	A theory is \emph{operationally valid} iff $\forall \phi \in {\gleadsto}(\allargs), \ar_0 \in \gargs[\phi], \ar_1 \in \allargs: (\ar_1, \ar_0) \ileadstoe ¬\phi ⇒ [\ar_1 \gbeats \ar_0] \land [\exists \ar_2 \in {\gbeatsinv}(\ar_1) \suchthat (\ar_0 + \ar_2, \ar_1) \ileadstoe \phi]$.
\end{definition}

\subsection{Example model}
\begin{itemize}
	\item $\ar_0 = \text{“Veg bec of ethics and health”}$
	\item $\ar_1 = \text{“Ethics is not defined”}$
	\item $\ar_2 = \text{“It is since Aristotle at least”}$
	\item $\ar_1' = \text{“Health isn’t important”}$
	\item $\ar_2' = \text{“Ask my grandmother”}$
	\item $\ar_1 \mbeats \ar_0$, $\ar_1' \mbeats \ar_0$
	\item $\ar_2 \mbeats \ar_1$, $\ar_2' \mbeats \ar_1'$
	\item $\ar_0 + \ar_2 \mbeats \ar_1$
	\item $\ar_1' \mbeats \ar_0 + \ar_2$
	\item $\ar_0 + \ar_2' \mbeats \ar_1'$
	\item $\ar_1 \mbeats \ar_0 + \ar_2'$
	\item $\ar_0 + \ar_2 + \ar_2' \mbeats \ar_1$
	\item $\ar_0 + \ar_2 + \ar_2' \mbeats \ar_1'$
\end{itemize}

\subsection{Theorem}
\begin{theorem}
	If a theory is answerable, has bounded length, and is operationally valid, it is valid.
\end{theorem}
\begin{proof}
	TODO adapt to new definitions!

	Given $\phi \in {\gleadsto}(\allargs), \args_2, \args_0 \subseteq \gargs$, define $\args_2$ as \emph{$\phi$-defending} $\args_0$ iff $\forall \ar_1 \in \allargs: [\exists \ar_0 \in \args_0 \suchthat (\ar_1, \ar_0) \ileadstoe ¬\phi] ⇒ [\exists \ar_2 \in \args_2 \suchthat (\ar_2, \ar_1) \ileadstoe \phi]$. Define $\args_2$ as \emph{decisively $\phi$-defending} $\args_0$ iff $\forall \ar_1 \in \allargs: [\exists \ar_0 \in \args_0 \suchthat (\ar_1, \ar_0) \ileadstoe ¬\phi] ⇒ [\exists \ar_2 \in \args_2 \suchthat (\ar_2, \ar_1) \ileadstost \phi]$.
